Post-Disturbance Tree Community Trajectories in a Neotropical Forest

### 2019-02-14 WholePlotTrajectories.Rmd
intro

An early conceptual basis of the linkage between biodiversity and disturbance is the Intermediate Disturbance Hypothesis (IDH).
The IDH assumes a relationship between community diversity and the intensity and frequency of disturbance events [@Connell1978].
The theory and states that community diversity peaks at intermediate level of disturbance.
The fluctuations of environmental conditions after disturbance are assumed to prevent competitive exclusion in fostering both competitively superior species and fast colonizers [@Shea2004;@Pulsford2016].
In highly diverse tropical forests, however, observations of the IDH are often blurry compared to theoretical expectations [@Hugues2007;@Sheil2003;@Norden2017], and the validation of the theory remains controversial [@Hubbell2001;@Fox2013;@Sheil2013].
Given the huge diversity of tropical forests and the variety of ecological processes that may underlie the IDH The scope of the IDH need be clarified regarding the different facets of community richness, abundance distribution and functioning.
The underlying processes of the IDH are besides part of the post-disturbance succession and the question is on the time after disturbance validating the hypothesis.
In this study we aimed as well to clarify the scope of the IDH and discuss its implication for the fate of tropical forests [@Sheil2003;@Shea2004].

>>>>

An early conceptual basis of the linkage between biodiversity and disturbance is the Intermediate Disturbance Hypothesis (IDH).
The IDH assumes a relationship between community diversity and the intensity and frequency of disturbance events [@Connell1978].
The theory and states that community diversity peaks at intermediate level of disturbance.
The fluctuations of environmental conditions after disturbance are assumed to prevent competitive exclusion in fostering both competitively superior species and fast colonizers [@Shea2004;@Pulsford2016].
In tropical forests, however, observations of the IDH are often blurry compared to theoretical expectations [@Hugues2007;@Sheil2003;@Norden2017], and the validation of the theory remains controversial [@Hubbell2001;@Fox2013;@Sheil2013].
The link between biodiversity and disturbance might be complicated by the huge diversity of tropical communities and by the variety of ecological processes that may underlie the IDH (colonisation, facilitation, coexistence,etc)[@Lindenmayer2012;@Garcia_florez2017].
The scope of the IDH need be clarified regarding the different facets of community diversity, like taxonomic richness, abundance distribution and functioning.
Besides, the question is highlighting the moment when emerge the conditions for the IDH, which would be an intermediate time after disturbance. 
In this study we then aimed to clarify the scope of the IDH in time and regarding the different facets of community diversity [@Sheil2003;@Shea2004].

### 2019-02-14 WholePlotTrajectories.tex
intro

The large areas covered with tropical forests worldwide hold crucial
environmental, economic and social values. They provide wood and
multiple non-timber forest products, shelter a diversified fauna, and
ensure cultural and human well-being. They regulate as well the local
and regional climates, and the carbon, water and nutrient cycles.
However, the growing demand in forests products together with current
global changes increase the pressure on remaining natural forests
\citep{Morales-Hidalgo2015}. These threats affect the natural
disturbance regime that shape and maintain the structure, composition
and functioning of tree communities
\citep{Schnitzer2001, Anderson-Teixeira2013, Sist2015}. It is now urgent
to understand the processes underlying community response to
disturbance, and thus to be able to anticipate the fate of tropical
forests in the current context. In tropical forests, disturbance
maintains a forest environment and initiates a suit of changes in the
structure and functioning of communities consituting a secondary
succession. Disturbance change both community abiotic and biotic
environments, through modifications in the fluxes of light, heat and
water \citep{Goulamoussene2017}. For now, post-disturbance trajectories
has been largely studied through forest structural parameters such as
aboveground biomass, tree height or stem density
\citep{Piponiot2016, Rutishauser2016}. Models based on observed
trajectories then assessed community time to recovery and identified
some of the determinants of post-disturbance trajectories, like forest
structure, forest composition, environmental parameters and disturbance
characteristics \citep{Herault2018}. Regarding community diversity and
composition, however, post-disturbance trajectories have not been as
thoroughly understood \citep{Guitet2018, Molino2001}. Given the variety
of species response to disturbance and the huge diversity of tropical
forests, biodiversity trajectories following disturbance might be very
diverse \citep{Lindenmayer2012, Garcia_florez2017}. We assumed that the
rate, endpoint and duration of the trajectories depend on the intensity
and frequency of disturbance: we then aimed to highlight the link
between biodiversity and disturbance and thus anticipate community
post-disturbance trajectories \citep{Chazdon2003a, Willig2018}.

An early conceptual basis of the linkage between biodiversity and
disturbance is the Intermediate Disturbance Hypothesis (IDH). The IDH
assumes a relationship between community diversity and the intensity and
frequency of disturbance events \citep{Connell1978}. The theory and
states that community diversity peaks at intermediate level of
disturbance. The fluctuations of environmental conditions after
disturbance are assumed to prevent competitive exclusion in fostering
both competitively superior species and fast colonizers
\citep{Shea2004, Pulsford2016}. In highly diverse tropical forests,
however, observations of the IDH are often blurry compared to
theoretical expectations \citep{Hugues2007, Sheil2003, Norden2017}, and
the validation of the theory remains controversial
\citep{Hubbell2001, Fox2013, Sheil2013}. The scope of the IDH should
then be clarified regarding the different facets of community richness,
abundance distribution and functioning. The underlying processes of the
IDH are besides part of the post-disturbance succession and the question
is on the time after disturbance validating the hypothesis. In this
study we aimed as well to clarify the scope of the IDH and discuss its
implication for the fate of tropical forests
\citep{Sheil2003, Shea2004}.

>>>>

The large areas covered with tropical forests worldwide hold crucial
environmental, economic and social values. They provide wood and
multiple non-timber forest products, shelter a diversified fauna, and
ensure cultural and human well-being. They regulate as well the local
and regional climates, and the carbon, water and nutrient cycles.
However, the growing demand in forests products together with current
global changes increase the pressure on remaining natural forests
\citep{Morales-Hidalgo2015}. These threats affect the natural
disturbance regime that shape and maintain the structure, composition
and functioning of tree communities
\citep{Schnitzer2001, Anderson-Teixeira2013, Sist2015}. It is urgent to
understand community response to disturbance and the processes
underlying to anticipate the fate of tropical forests in the current
context. Disturbance change both community abiotic and biotic
environments, through modifications in the fluxes of light, heat and
water \citep{Goulamoussene2017}. In tropical forests, the forest
environment is maintained after disturbance and there is a suit of
changes in the structure and functioning of communities that consitutes
a secondary succession. For now, this succession has been largely
studied through post-disturbance trajectories of forest structural
parameters such as aboveground biomass, tree height or stem density
\citep{Piponiot2016, Rutishauser2016}. Models based on observed
trajectories then assessed community time to recovery and identified
some of the determinants of post-disturbance trajectories, like forest
structure, forest composition, environmental parameters and disturbance
characteristics \citep{Herault2018}. Regarding community diversity and
composition, however, post-disturbance trajectories have not been as
thoroughly understood \citep{Guitet2018, Molino2001}. Given the variety
of species response to disturbance and the huge diversity of tropical
forests, post-disturbance biodiversity trajectories might be very
diverse \citep{Lindenmayer2012, Garcia_florez2017}. It is essential to
anticipate these trajectories, and hence to highlight the link between
biodiversity and disturbance, specifically how disturbance intensity and
frequency determine the rate, endpoint and duration of the trajectories
\citep{Chazdon2003a, Willig2018}.

An early conceptual basis of the linkage between biodiversity and
disturbance is the Intermediate Disturbance Hypothesis (IDH). The IDH
assumes a relationship between community diversity and the intensity and
frequency of disturbance events \citep{Connell1978}. The theory and
states that community diversity peaks at intermediate level of
disturbance. The fluctuations of environmental conditions after
disturbance are assumed to prevent competitive exclusion in fostering
both competitively superior species and fast colonizers
\citep{Shea2004, Pulsford2016}. In tropical forests, however,
observations of the IDH are often blurry compared to theoretical
expectations \citep{Hugues2007, Sheil2003, Norden2017}, and the
validation of the theory remains controversial
\citep{Hubbell2001, Fox2013, Sheil2013}. The link between biodiversity
and disturbance might be complicated by the huge diversity of tropical
communities and by the variety of ecological processes that may underlie
the IDH (colonisation, facilitation,
coexistence,etc)\citep{Lindenmayer2012, Garcia_florez2017}. The scope of
the IDH need be clarified regarding the different facets of community
diversity, like taxonomic richness, abundance distribution and
functioning. Besides, the question is highlighting the moment when
emerge the conditions for the IDH, which would be an intermediate time
after disturbance. In this study we then aimed to clarify the scope of
the IDH in time and regarding the different facets of community
diversity \citep{Sheil2003, Shea2004}.
